---
title: "p8105_hw1_qc2337"
output: github_document
date: "2024-09-21"
---
# hw1 Problem 1

First loaded the palmerpenguins package and load pemnguins dataset

```{r}

library(palmerpenguins)
data("penguins", package = "palmerpenguins")

```

Second description of the penguins dataset using inline R code
show variable names, structure of the dataset, size of the dataset, mean flipper length calculation

```{r}

names(penguins)

str(penguins)

nrow(penguins)
ncol(penguins)

mean_flipper_length <- mean(penguins$flipper_length_mm, na.rm = TRUE)
mean_flipper_length

```

From the description code, we know that in the penguins dataset,
variable names include: `species`, `island`, `bill_length_mm`,`bill_depth_mm`,  `flipper_length_mm`, `body_mass_g`, `sex`, `year`.

The size of the dataset includes: 344 rows and 8 columns.

The mean flipper length is 200.9152.


Make a scatterplot and save the scatterplot

```{r}

library(ggplot2)
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
    geom_point() + 
    labs(title = "Scatterplot of Flipper Length vs Bill Length by Species",
         x = "Bill Length (mm)",
         y = "Flipper Length (mm)")

ggsave("penguins_scatterplot.png")

```

# hw1 Problem 2
This is code for the problem2 at P8105.

Loaded necessary packages (e.g. `tidyverse`) for problem 2.

```{r}
library(tidyverse)
```

Created `p2_df` as the data frame, comprised of:

* a random sample of size 10 from a standard Normal distribution

* a logical vector indicating whether elements of the sample are greater than 0

* a character vector of length 10

* a factor vector of length 10, with 3 different factor “levels”

```{r}
  p2_df = 
    tibble(
      norm_samp = rnorm(n = 10),
      samp_g0 = norm_samp > 0,
      char_samp = letters [1:10],
      fact_samp = factor(rep(c("1", "2", "3"), length = 10))
    )
print(p2_df)
```

Take the mean of each variable in the `p2_df`

```{r}
  mean_norm_samp = mean(pull(p2_df, norm_samp), na.rm = TRUE)
  mean_norm_samp 

  mean_samp_g0 =  mean(pull(p2_df, samp_g0), na.rm = TRUE)
  mean_samp_g0

  mean_char_samp = mean(pull(p2_df, char_samp), na.rm = TRUE)
  mean_char_samp
  
  mean_fact_samp = mean(pull(p2_df, fact_samp), na.rm = TRUE)
  mean_fact_samp
```

I found that the `random sample of size 10`  and `logical vector ` works,
  but `character vector` and `factor vector ` doesn't work.
  
Applies the `as.numeric` function to the logical, character, and factor variables

```{r echo=TRUE, eval=FALSE}

numeric_samp_g0 = as.numeric(pull(p2_df, samp_g0))

numeric_char_samp = as.numeric(pull(p2_df, char_samp))

numeric_fact_samp = as.numeric(pull(p2_df, fact_samp))

```

I find that:

The logical variable turns into 2 numeric: `TRUE` becomes `1` and `FALSE` becomes `0`.

The character variable turns to `NA` for all elements, R cannot calculate the mean. So when I try to take the mean of character variable, it shows warning message: argument is not numeric or logical: returning NA.

The factor variable turns to numeric variable which is the factor levels internally in R. Now I can take the mean of factor variable, because my factor variable is in order. 


